\documentclass[../midgard.tex]{subfiles}
\graphicspath{{\subfix{../images/}}}
\begin{document}

\section*{Introduction}
\label{h:introduction}
\addcontentsline{toc}{chapter}{Introduction}

Midgard is a Layer 2 (L2) scaling solution for the Cardano blockchain.
It employs optimistic rollup technology to enhance Cardano's capacity to process transactions and host more complex applications, delivering a richer user experience at a more competitive cost.
As Cardano continues to grow in usage and demand, scaling solutions like Midgard are critical for maintaining high performance and low transaction costs.
This whitepaper describes the architecture and technical design of the Midgard protocol, detailing how it integrates with Cardano's Layer 1 (L1) to deliver secure and efficient transaction processing.

Optimistic rollups process blocks of transactions off-chain and commit those blocks' headers onchain to the L1 ledger.
Each block is committed by a Midgard operator who guarantees the block's validity.
The block then waits in a queue for at least a fixed duration to be merged into the confirmed state of the optimistic rollup on L1.
The operator must collateralize their guarantee with a bond deposit and publish the full contents of the block on the publicly accessible data availability (DA) layer.

While a committed block is queued, anyone can inspect its contents on the data availability layer and ascertain whether it is valid.
If someone detects that the block is invalid, they can submit a fraud proof to prevent it from being merged into the confirmed state, slash the operator's bond, and receive a portion of the forfeited bond as a reward.
Thus, an optimistic rollup can process a large number of transactions offchain while maintaining security and finality properties that are similar to transactions processed directly onchain, as long as:
\begin{itemize}
    \item The bond requirement for the rollup's operators is large enough to discourage fraud.
    \item The reward for preventing an invalid block from merging is large enough to encourage public vigilance in watching the operators.
    \item The waiting period for committed blocks is long enough to allow the watchers to detect and prove fraud before those blocks are merged.
    \item The data availability layer is accessible by anyone who wishes to inspect the rollup blocks at any time that they wish to do so.
\end{itemize}

Whenever the latter three security parameter values are calibrated to provide a high probability of invalid blocks being detected and disqualified, the bond requirement is a strong deterrent against operators attempting fraud.
An operator cannot dismiss the forfeited bond as merely a ``cost of doing business'' paid to obtain potentially larger revenues from fraud.
Whenever an invalid block is disqualified, it does \emph{not} affect the confirmed state, so there are no revenues from that fraud to offset the operator's forfeited bond.

The main design goal of Midgard is to streamline the processes by which blocks are committed/merged, fraud is detected, and fraud proofs are verified onchain.
Advancing this goal allows the security parameters to be calibrated to achieve a better balance between security, transaction throughput, confirmation time, and community participation in committing blocks and detecting fraud.

\subsection*{Scalability and efficiency}
\label{h:scalability-and-efficiency}

By processing transactions off-chain and only validating them on-chain when fraud proofs challenge them, Midgard significantly increases throughput and reduces costs for Cardano transactions.
Its rollup blocks use sparse Merkle trees and compact state representations to enhance the protocol's efficiency further, enabling it to handle a large volume of transactions without overburdening the L1.

The deterministic nature of Cardano transactions allows Midgard fraud proofs to pinpoint the specific site of a transaction that violated Midgard's ledger rules, without having to look at any other parts of that transaction, any other unrelated transactions within the block, or any other blocks.
This keeps fraud proofs and their onchain validation procedures small and efficient, which reduces the time and cost needed to submit fraud proofs when invalid blocks are detected, which makes it feasible for a wider group of people to police Midgard's blocks.
In this way, Midgard significantly reduces fraud proof size relative to optimistic rollups used in Ethereum and other account-based blockchain ecosystems, where a much larger part of the global blockchain state needs to be inspected when constructing and verifying a fraud proof.

\subsection*{Censorship resistance and fallback mechanisms}

On its own, the optimistic rollup mechanism described above ensures a high-level of assurance for the validity of block headers committed to the state queue and merged to Midgard's confirmed state.
However, it does not prevent operators from censoring users' deposits, withdrawals, and L2 transactions.
Consequently, Midgard's consensus protocol includes additional smart contract mechanisms to provide censorship resistance for these events.

Midgard deposits and withdrawals are initiated via L1 smart contracts that assign definite inclusion times to them.
An operator block is invalid if it contains these inclusion times in its event interval but fails to include the associated deposit or withdrawal events.
This ensures that if operators continue committing blocks to Midgard's state queue, then they cannot ignore deposit and withdrawal events.

Midgard L2 transaction requests are typically submitted to operators via a publicly accessible API, and they can be ignored by operators.
However, any user can escalate his L2 transaction request by posting a transaction order on L1.
Similar to Midgard deposits and withdrawals, an L1 transaction order is assigned an inclusion time that guarantees its inclusion in a subsequent valid block.

If Midgard operators stop committing blocks at all to the state queue, then the inclusion times on their own cannot guarantee that deposits, withdrawals, and L2 transactions will be processed in a timely manner.
However, for this extreme case, Midgard's consensus protocol includes the escape hatch mechanism, which temporarily lowers the bond requirement for new operators.
This makes it easier for honest actors to become Midgard operators and start committing blocks to the state queue.
This ensures that user funds cannot be stranded on Midgard even if its current operators entirely stop committing blocks.

\subsection*{Feedback and contributions welcome!}
\label{h:feedback-and-contributions-welcome}

This document specifies the ledger, onchain contracts, and offchain components of the Midgard architecture.
We detail the roles of users, operators, and watchers, how blocks are committed and merged on L1, and how fraud proofs are handled.

We invite developers, validators, and the broader community to engage with this document and offer feedback and contributions to help refine and enhance the protocol.

\end{document}
