\documentclass[../midgard.tex]{subfiles}
\graphicspath{{\subfix{../images/}}}
\begin{document}

\section{State queue}
\label{h:state-queue}

The state queue is an L1 data structure that stores Midgard operators' committed block headers until they are confirmed.
Active operators from the operator directory (see \cref{h:operator-directory}) take turns committing block headers to the state queue according to the rotating schedule enforced by the scheduler (see \cref{h:scheduler}).

\subsection{Utxo representation}
\label{h:state-queue-utxo-representation}

The \code{state\_queue} is implemented as a key-unordered linked list of block headers (see \cref{h:block}).
Each \code{state\_queue} node's key is its block header's hash.
\begin{align*}
    \T{StateQueueDatum} &\coloneq \T{NodeDatum}(\T{Header}) \\ \\
    \T{valid\_key}(\T{scd} : \T{StateQueueDatum}) &\coloneq
        \Bigl( \T{scd.key} \equiv \T{Some}(\T{hash}(\T{scd.data})) \Bigr)
\end{align*}

Committing a block header to the \code{state\_queue} means appending a node containing the block header to the end of the queue.
After staying there for the \code{maturity\_duration} (a protocol parameter), it is merged to the confirmed state (held at the \code{state\_queue} root node) in the first-in-first-out (FIFO) order.

\subsection{Minting policy}
\label{h:state-queue-minting-policy}

The \code{state\_queue} minting policy controls the structural changes to the state queue.
It is statically parametrized on the \code{hub\_oracle}, \code{active\_operators}, \code{retired\_operators}, \code{scheduler}, and \code{fraud\_proof} minting policies.
Redeemers:

\begin{description}
    \item[Init.] Initialize the \code{state\_queue} via the Midgard hub oracle.
      Conditions:
        \begin{enumerate}
            \item The transaction must mint the Midgard hub oracle token.
            \item The transaction must Init the \code{state\_queue}.
            \item The validity range of the transaction must be bounded on both ends, and must not be longer than five minutes (TODO).
            \item The inclusive upper bound of the validity range must be used as both the \code{start\_time} and \code{end\_time} of the produced confirmed state data.
            \item \code{header\_hash} and \code{prev\_header\_hash} must have genesis values (28 zero bytes).
            \item \code{utxo\_root} must be set to the root of an empty Merkle tree.
            \item \code{protocol\_version} must be set to 0 (TODO).
        \end{enumerate}

    \item[Commit Block Header.] An operator commits a block header to the state queue if it is the operator's turn according to the rotating schedule.
      Grouped conditions:
        \begin{itemize}
            \item Commit the block header to the state queue, with the operator's consent:
            \begin{enumerate}
                \item Let \code{operator} be a redeemer argument indicating the key of the operator committing the block header.
                \item The transaction must be signed by \code{operator}.
                \item The transaction must Append a node to the \code{state\_queue}.
                  Among the transaction outputs, let that node be \code{header\_node} and its predecessor node be \code{previous\_element} (this can be either the confirmed state, or a previous block).
                \item \code{operator} must be the operator of \code{header\_node}.
                \item The key of \code{header\_node} must be the hash of its underlying \code{Node} data.
            \end{enumerate}

            \item Verify that it is the operator's turn to commit according to the scheduler:
            \begin{enumerate}[resume]
                \item The transaction must include a reference input with the scheduler NFT.
                  Let that input be the \code{scheduler\_state}.
                \item The \code{operator} field of \code{scheduler\_state} must match \code{operator}.
            \end{enumerate}

            \item Verify block timestamps:
            \begin{enumerate}[resume]
                \item The \code{start\_time} of \code{header\_node} must be equal to the \code{end\_time} of the \code{previous\_element}.
                \item The \code{end\_time} of \code{header\_node} must match the transaction's time-validity upper bound exclusively (TODO).
                \item The \code{end\_time} of \code{header\_node} must be within the shift interval of \code{scheduler\_state}.
            \end{enumerate}
            
            \item Update the operator's timestamp in the active operators set:
            \begin{enumerate}[resume]
                \item The transaction must include an input, spent via the Update Bond Hold New State redeemer, of an \code{active\_operators} node with a key matching the \code{operator}.
            \end{enumerate}
            \item New block must carry over data from \code{previous\_element} properly:
                \begin{enumerate}[resume]
                  \item Its \code{header\_hash} must be preserved under new node's \code{prev\_header\_hash}.
                  \item Its ledger must be preserved under new node's \code{prev\_utxos\_root}.
                  \item Protocol version must also be preserved from \code{previous\_element} (TODO).
                \end{enumerate}
        \end{itemize}

    \item[Merge To Confirmed State.] If a block header is mature, merge it to the confirmed state of the \code{state\_queue}.
      Conditions:
    \begin{enumerate}
      \item The transaction must Fold the first node of \code{state\_queue} into its root.
        Let \code{header\_node} be the folded node, \code{confirmed\_state\_node\_input} be its predecessor node before removal, and \code{confirmed\_state\_node\_output} be the remaining node after removal.
        \item \code{confirmed\_state\_node\_input} and \code{confirmed\_state\_node\_output} must both be root nodes of \code{state\_queue}.
        \item \code{header\_node} must be mature---the lower bound of the transaction validity interval meets or exceeds the sum of the \code{end\_time} field of \code{header\_node} and the Midgard \code{maturity\_duration} protocol parameter.
        \item \code{confirmed\_state\_node\_output} must match:
          \begin{enumerate}
            \item \code{start\_time} and \code{header\_hash} from \code{confirmed\_state\_node\_input} under \code{start\_time} and \code{prev\_header\_hash} respectively.
            \item Key of \code{header\_node} under \code{header\_hash}, along with its ledger and protocol version (TODO?) under \code{utxo\_root} and \code{protocol\_version} respectively.
          \end{enumerate}
        \item If either the \code{deposits\_root}, \code{withdrawals\_root} or \code{transactions\_root} of \code{header\_node} is \emph{not} the MPT root hash of the empty set, a \code{settlement} utxo must be produced via the Spawn redeemer.
          The redeemer must mention \code{settlement\_id} equal to \code{header\_node} as a coupling medium.
        \item Otherwise no settlement minting must be present in the transaction.
    \end{enumerate}

    \item[Remove Fraudulent Block Header.] Remove a fraudulent block header or its link from the state queue and slash its operator's ADA bond.
      Grouped conditions:
    \begin{itemize}
        \item Slash the fraudulent operator.
          Let \code{fraudulent\_operator} be a redeemer argument indicating the operator who committed the fraudulent block header:
        \begin{enumerate}
            \item If the operator is active:
                \begin{enumerate}
                    \item The transaction must Remove a node from the \code{active\_operators} set via the Remove Operator Bad State redeemer.
                      The \code{slashed\_operator} argument provided to that redeemer must match \code{fraudulent\_operator}.
                \end{enumerate}
            \item If the operator is retired:
                \begin{enumerate}
                    \item The transaction must Remove a node from the \code{retired\_operators} set via the Remove Operator Bad State redeemer.
                      The \code{slashed\_operator} argument provided to that redeemer must match \code{fraudulent\_operator}.
                \end{enumerate}
            \item Otherwise, if the operator is already slashed:
                \begin{enumerate}
                    \item The transaction must refer to elements from both the \code{active\_operators} and \code{retired\_operators} sets and show that neither lists include the fraudulent operator.
                \end{enumerate}
        \end{enumerate}

        \item Let \code{fraudulent\_block\_header\_hash} be a redeemer argument specifying the header hash of the fraudulent block with a minted proof token.
        \item Remove the fraudulent block header or its link.
          Two routes are allowed:
          \begin{itemize}
            \item Fraudulent block has a link and that link node is being removed.
              \begin{enumerate}
                \item Let \code{fraudulent\_node} be the anchor node, and \code{removed\_node} be its child node that is being removed.
                \item Key of \code{fraudulent\_node} must be equal to \code{fraudulent\_block\_header\_hash}.
                \item \code{operator\_vkey} from the \code{removed\_node} must match \code{fraudulent\_operator}.
              \end{enumerate}
            \item Fraudulent block is the last node in state queue and is being removed.
              \begin{enumerate}
                \item Let \code{fraudulent\_node} be the removing node, and \code{anchor\_element} be its parent element.
                \item Key of \code{fraudulent\_node} must be equal to \code{fraudulent\_block\_header\_hash}.
                \item \code{operator\_vkey} from the \code{fraudulent\_node} must match \code{fraudulent\_operator}.
                \item \code{fraudulent\_node} must not link to a child node.
              \end{enumerate}
          \end{itemize}

        \item Verify that fraud has been proved for the removed node or its predecessor:
        \begin{enumerate}
            \item The transaction must include a reference input holding a \code{fraud\_proof} token.
            \item Let \code{fraud\_proof\_block\_hash} be the last 28 bytes of the \code{fraud\_proof} token.
            \item \code{fraud\_proof\_block\_hash} must match \code{fraudulent\_block\_header\_hash} from the redeemer.
        \end{enumerate}
    \end{itemize}
\end{description}

\subsection{Spending validator}
\label{h:state-queue-spending-validator}

The spending validator of \code{state\_queue\_addr} always forwards to its corresponding minting policy (statically parametrized) and requires the transaction to invoke it.
It does not allow any in-place modifications to the \code{data} field of nodes in \code{state\_queue}.
Conditions:
\begin{enumerate}
    \item The transaction must mint or burn tokens of the \code{state\_queue} minting policy.
\end{enumerate}

\end{document}
